Memory for television episodes preserves event content while introducing new across-event similarities

### 2020-02-18 paper/main.tex
fixed another typo

\title{Memory for television episodes preserves event content while introducing new across-event similarities}

@@

\begin{abstract}
The ways our experiences unfold over time define unique \textit{trajectories} through the relevant representational spaces.  Within this geometric framework, one can compare the shape of the trajectory formed by an experience to that defined by our later remembering of that experience.  We propose a framework for mapping naturalistic experiences onto geometric spaces that characterize how experiences are segmented into discrete events, and how the contents of a event sequences evolve over time.  We apply this approach to a naturalistic memory experiment which had participants view and recount a television episode.  The content of participants' recountings of events from the original episode closely matched the original episode's content.  However, the similarity patterns \textit{across} events was much different in the original episode as compared with participants' recountings. We also identified a network of brain structures that are sensitive to the ``shapes'' of ongoing experiences, and an overlapping network that is sensitive (at the time of encoding) to how people later remembered those experiences in relation to other experiences.  In this way, modeling the content of richly structured experiences can reveal how (geometrically and conceptually) those experiences are segmented into events and integrated into our memories of other experiences.
\end{abstract}

@@

Occasionally, this gradual ``drift" of our ongoing experience is punctuated by sudden changes, or ``shifts"~\citep[e.g., when we walk through a doorway; ][]{RadvZack17}.  Prior research suggests that these sharp transitions (termed \textit{event boundaries}) help to discretize our experiences (and their mental representations) into \textit{events}~\citep{RadvZack17, BrunEtal18, HeusEtal18b, ClewDava17, EzzyDava11, DuBrDava13}.  The interplay between the stable (within-event) and transient (across-event) temporal dynamics of an experience also provides a potential framework for transforming experiences into memories that distill those experiences down to their essence.  For example, prior work has shown that event boundaries can influence how we learn sequences of items~\citep{HeusEtal18b, DuBrDava13}, navigate~\citep{BrunEtal18}, and remember and understand narratives~\citep{ZwaaRadv98, EzzyDava11}. Prior research has implicated the hippocampus and the medial prefrontal cortex as playing a critical role in transforming experiences into stuctured and consolidated memories ~\citep{TompDava17}.

@@

\section*{Results}
To characterize the ``essence" of the \textit{Sherlock} episode and participants' subsequent recountings of its unfolding, we used a topic model~\citep{BleiEtal03} to discover the latent themes in the episode's dynamic content.  Topic models take as inputs a vocabulary of words to consider and a collection of text documents, and return two output matrices.  The first of these is a \textit{topics matrix} whose rows are topics (latent themes) and whose columns correspond to words in the vocabulary. The entries of the topics matrix define how each word in the vocabulary is weighted by each discovered topic.  For example, a detective-themed topic might weight heavily on words like ``crime,'' and ``search.''  The second output is a \textit{topic proportions matrix}, with one row per document and one column per topic.  The topic proportions matrix describes what mixture of discovered topics is reflected in each document.

\cite{ChenEtal17} collected hand-annotated information about each of 1000 (manually identified) time segments spanning the roughly 50 minute video used in their experiment.  This information included: a brief narrative description of what was happening, the location where the scene took place, the names of any characters on the screen, and other similar details (for a full list of annotated features, see \textit{Methods}).  We took from these annotations the union of all unique words (excluding stop words, such as ``and," ``or," ``but," etc.) across all features and scenes as the ``vocabulary" for the topic model.  We then concatenated the sets of words across all features contained in overlapping, sliding windows of (up to) 50 scenes, and treated each window as a single ``document" for the purpose of fitting the topic model.  Next, we fit a topic model with (up to) $K = 100$ topics to this collection of documents.  We found that 32 unique topics (with non-zero weights) were sufficient to describe the time-varying content of the video (see \textit{Methods}; Figs.~\ref{fig:schematic}, \topics).  Note that our approach is similar in some respects to Dynamic Topic Models~\citep{BleiLaff06} in that we sought to characterize how the thematic content of the episode evolved over time.  However, whereas Dynamic Topic Models are designed to characterize how the properties of \textit{collections} of documents change over time, our sliding window approach allows us to examine the topic dynamics within a single document (or video).  Specifically, our approach yielded (via the topic proportions matrix) a single \textit{topic vector} for each sliding window of annotations transformed by the topic model.  We then stretched the resulting windows-by-topics matrix to match the time series of the 1976 fMRI volumes collected as participants viewed the episode.

\begin{figure}[tp]
\centering
\includegraphics[width=1\textwidth]{figs/schematic}
\caption{\small \textbf{Methods overview.} We used hand-annotated descriptions of each moment of video to fit a topic model.  Three example video frames and their associated descriptions are displayed (top two rows).  Participants later recalled the video (in the third row, we show example recalls of the same three scenes from participant 13).  We used the topic model (fit to the annotations) to estimate topic vectors for each moment of video and each sentence the participants recalled.  Example topic vectors are displayed in the bottom row (blue: video annotations; green: example participant's recalls).  Three topic dimensions are shown (the highest-weighted topics for each of the three example scenes, respectively).  We also show the 10 highest-weighted words for each topic.  Figure~\topics~provides a full list of the top 10 words from each of the discovered topics.}
\label{fig:schematic}
\end{figure}
% Odd name for figure in "Results" section rather than "methods" section 
% - Paxton

@@

\begin{figure}[tp]
\centering
\includegraphics[width=\textwidth]{figs/eventseg}
\caption{\small \textbf{Modelling naturalistic stimuli and recalls.} All panels: darker colors indicate greater values; range: [0, 1].  \textbf{A.} Topic vectors ($K = 100$) for each of the 1976 video timepoints.  \textbf{B.} Timepoint-by-timepoint correlation matrix of the topic vectors displayed in Panel A.  Event boundaries discovered by the HMM are denoted in yellow (30 events detected).  \textbf{C.} Average topic vectors for each of the 30 video events. \textbf{D.} Topic vectors for each of 265 sliding windows of sentences spoken by an example participant while recalling the video.  \textbf{E.} Timepoint-by-timepoint correlation matrix of the topic vectors displayed in Panel D. Event boundaries detected by the HMM are denoted in yellow (22 events detected).  For similar plots for all participants see Figure~\corrmats.  \textbf{F.} Average topic vectors for each of the 22 recalled events from the example participant.  \textbf{G.} Correlations between the topic vectors for every pair of video events (Panel C) and recalled events (from the example participant; Panel F).  For similar plots for all participants, see Figure~\matchmats.  \textbf{H.} Average correlations between each pair of video events and recalled events (across all 17 participants).  To create the figure, each recalled event was assigned to the video event with the most correlated topic vector (yellow boxes in panels G and H).  The heat maps in each panel were created using \texttt{Seaborn}~\citep{WaskEtal16}.}
\label{fig:model}
\end{figure}
% I'm all for citing software packages, but where what's the distinction between packages we do cite (seaborn, brainiak, sklearn, hypertools, quail, umap, etc.) and don't cite (wordcloud, nilearn, fastdtw, etc.)?
% - Paxton

@@

Although the example participant's recall topic proportions matrix has some visual similarity to the video topic proportions matrix, the time-varying topic proportions for the example participant's recalls are not as sparse as those for the video (compare Figs.~\ref{fig:model}A and D).  Similarly, although there do appear to be periods of stability in the recall topic dynamics (i.e., most topics are active or inactive over contiguous blocks of time), the individual topics' overall timecourses are not as cleanly delineated as the video topics'.  To examine these patterns in detail, we computed the timepoint-by-timepoint correlation matrix for the example participant's recall topic trajectory (Fig.~\ref{fig:model}E).  As in the video correlation matrix (Fig.~\ref{fig:model}B), the example participant's recall correlation matrix has a strong block diagonal structure, indicating that their recalls are discretized into separated events.  As for the video correlation matrix, we can use an HMM, along with the aforementioned number-of-events optimization procedure (also see \textit{Methods}) to determine how many events are reflected in the participant's recalls and where specifically the event boundaries fall (outlined in yellow).  We carried out a similar analysis on all 17 participants' recall topic proportions matrices (Fig.~\corrmats).

Two clear patterns emerged from this set of analyses.  First, although every individual participant's recalls could be segmented into discrete events (i.e., every individual participant's recall correlation matrix exhibited clear block diagonal structure; Fig.~\corrmats), each participant appeared to have a unique \textit{recall resolution}, reflected in the sizes of those blocks.  While, some participants' recall topic proportions segmented into just a few events (e.g., Participants P4, P5, and P7), others' segmented into many shorter duration events (e.g., Participants P12, P13, and P17).  This suggests that different participants may be recalling the video with different levels of detail-- e.g., some might touch on just the major plot points, whereas others might attempt to recall every minor scene or action.  The second clear pattern present in every individual participant's recall correlation matrix is that, unlike in the video correlation matrix, there are substantial off-diagonal correlations.  Whereas each event in the original video was (largely) separable from the others (Fig.~\ref{fig:model}B), in transforming those separable events into memory, participants appear to be integrating across multiple events, blending elements of previously recalled and not-yet-recalled content into each newly recalled event~\citep[Figs.~\ref{fig:model}D, \corrmats; also see][]{MannEtal11, HowaEtal12}.

The above results indicate that both the structure of the original video and participants' recalls of the video exhibit event boundaries that can be identified automatically by characterizing the dynamic content using a shared topic model and segmenting the content into events via HMMs.  Next, we asked whether some correspondence might be made between the specific content of the events the participants experienced in the video, and the events they later recalled.  One approach to linking the experienced (video) and recalled events is to label each recalled event as matching the video event with the most similar (i.e., most highly correlated) topic vector (Figs.~\ref{fig:model}G, \matchmats).  This yields a sequence of ``presented" events from the original video, and a (potentially differently ordered) sequence of ``recalled" events for each participant.  Analogous to classic list-learning studies, we can then examine participants' recall sequences by asking which events they tended to recall first~\citep[probability of first recall; Fig.~\ref{fig:list-learning}A;][]{AtkiShif68, PostPhil65, WelcBurn24}; how participants most often transition between recalls of the events as a function of the temporal distance between them~\citep[lag-conditional response probability; Fig.~\ref{fig:list-learning}B;][]{Kaha96}; and which events they were likely to remember overall~\citep[serial position recall analyses; Fig.~\ref{fig:list-learning}C;][]{Murd62a}. Interestingly, for two of these analyses (probability of first recall and lag-conditional response probability curves) we observe patterns comparable to classic effects from the list-learning literature: namely, a higher probability of initiating recall with the first event in the sequence (Fig.~\ref{fig:list-learning}A) and a higher probability of transitioning to neighboring events with an asymmetric forward bias (Fig.~\ref{fig:list-learning}C). In contrast, we do not observe a pattern comparable to the serial position effect (Fig.~\ref{fig:list-learning}C), but rather we see higher memory for specific events distributed somewhat evenly throughout the video.

We can also apply two list-learning-native analyses that describe how participants group items in their recall sequences: temporal clustering and semantic clustering \citep[][see \textit{Methods} for details]{PolyEtal09}.  Temporal clustering refers to the extent to which participants group their recall responses according to encoding position.  Overall, we found that sequentially viewed video events were clustered heavily in participants' recall event sequences (mean: 0.767, SEM: 0.029), and that participants with higher temporal clustering scores tended to perform better according to both \cite{ChenEtal17}'s hand-annotated memory scores (Pearson's $r(15) = 0.62,~p = 0.008$) and our model's estimate (Pearson's $r(15) = 0.54,~p = 0.024$).  Semantic clustering measures the extent to which participants cluster their recall responses according to semantic similarity.  We found that participants tended to recall semantically similar video events together (mean: 0.787, SEM: 0.018), and that semantic clustering score was also related to both hand-annotated  (Pearson's $r(15) = 0.65,~p = 0.004$) and model-derived (Pearson's $r(15) = 0.63,~p = 0.007$) memory performance.
% NOTE: 
% Temporal and semantic clustering are both positively correlated with n_events -- r=0.49; p=0.046 and r=0.61; p=0.009, respectively.  
% - Paxton

% NOTE: Pruning out to tone down the list-learning slander
%In list-learning studies, this set of three analyses may be used to gain a nearly complete view into the sequences of recalls participants made~\citep[e.g., ][]{Kaha12}.  Extending these analyses to apply to naturalistic stimuli and recall~\citep{HeusEtal17b} highlights that, in naturalistic recall, these analyses provide a wholly incomplete picture: they leave out any attempt to quantify participants' abilities to capture the \textit{content} of what occurred in the video-- their only experimental instruction!

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=1\textwidth]{figs/precision_distinctiveness}
  \caption{\small \textbf{Novel content-based metrics of naturalistic memory: precision and distinctiveness.} \textbf{A.} The video-recall correlation matrix for a representative participant (17).  The yellow boxes highlight the maximum correlation in each column.  The example participant's overall precision score was computed as the average across correlation values in the yellow boxes.  Their distinctiveness score was computed as the the average (over recall events) of 1 minus the average correlation between each recall event and all other recall events that do not display a box in the same row.  \textbf{B.} The (Pearson's) correlation between precision and hand-annotated memory performance. \textbf{C.} The correlation between distinctiveness and hand-annotated memory performance. \textbf{D.} The correlation between precision and the number of video events successfully recalled, as determined by our model. \textbf{E.} The correlation between distinctiveness and the number of video events successfully recalled, as determined by our model.}
  \label{fig:precision-distinctiveness}
\end{figure}

Statistical models of memory studies often treat recall success as binary (i.e., an item either was or was not recalled), or occasionally categorical \citep[e.g., to distinguish familiarity from recollection;][]{YoneEtal02}.  Such approaches are tenable in classical list-learning or recognition memory paradigms, as the presented stimuli tend to be very simple (e.g., a sequence of individual words or items).  However, the feature-rich content of a naturalistic experiences may later be described with many, highly variable levels of success.  Our framework produces a content-based model of individual stimulus and recall events by projecting the dynamic content of the video and participants' recalls into a shared topic space.  This allows for direct, quantitative comparison between all stimulus and recall events, as well as between the recall events themselves.  Leveraging these content-based models of the stimulus/recall events, we developed two novel, \textit{continuous} metrics for analyzing naturalistic memory:~\textit{precision} and \textit{distinctiveness}.  We define precision as the ``completeness" of recall, or how fully the presented content was recapitulated in memory.  Under our framework, we quantify this for a given recall event as the correlation between the topic proportions of the recall event and the maximally correlated video event (Fig.~\ref{fig:precision-distinctiveness}).  A second novel metric we introduce here is \textit{distinctiveness}, which we define as the ``specificity" of recall, or how unique the description of a given section of content was, compared to descriptions for other sections of content.  We quantify this for each recall event as 1 minus the average correlation between the given recall event and all other recall events not matched to the same video event.  In addition to individual events, one may also use these metrics to describe each participant's overall performance (i.e., by averaging across a participant's event-wise precision or distinctiveness scores).  Participants whose recall events are more veridical descriptions of what happened in the video event will presumably have higher precision scores. We find that, across participants, a higher precision score is correlated to both hand-annotated memory performance (Pearson's $r(15) = 0.60, p = 0.010$) and the number of video events successfully remembered, as determined by our model (Pearson's $r(15) = 0.90, p < 0.001$).  We also hypothesized that participants who recounted events in a more distinctive way would display better overall memory.  We find that this distinctiveness score is related to our model's estimated number of recalled events (Pearson's $r(15) = 0.55, p = 0.021$), and while we do not find distinctiveness to be related to hand-annotated memory performance (Pearson's $r(15) = 0.28, p = 0.275$), this is not entirely surprising given how the hand-annotated memory scores were computed (see \textit{Discussion}).

@@

Figure \ref{fig:distinctiveness-detail} similarly contrasts two example participants' recalls for a common video event (event 19) to illustrate the tangible differences between high and low distinctiveness scores.  Here, we have extracted the full set of sentences comprising the most distinctive recall event (P13) and least distinctive recall event (P11) recall event matched to the example video event (Fig.~\ref{fig:distinctiveness-detail}C).  We also extracted the annotations for the example video event, as well as those from each other video event whose content the example participants' single recall events described (Fig.~\ref{fig:distinctiveness-detail}B).  We then shaded the annotation text for each video event with a different color, and shaded each word of the example participants' recall text by the color of the video event it describes.  The majority of the most distinctive recall event text describes video event 19's content, with the first five and last one sentence describing the video events immediately preceding and succeeding the current one, respectively.  Meanwhile, the least precise participant's recall for video event 19 blends the content from five separate video events, does not transition between them in order, and often combines descriptions of two video events' content in the same sentence

The prior analyses leverage the correspondence between the 100-dimensional topic proportion matrices for the video and participants' recalls to characterize recall.  However, it is difficult to gain deep insights into the content of (or relationships between) experiences and memories solely by examining these topic proportions (e.g., Figs.~\ref{fig:model}A, D) or the corresponding correlation matrices (Figs.~\ref{fig:model}B, E, \corrmats).  And while we can directly examine the original text underlying these topic vectors (e.g., Figs.~\ref{fig:precision-detail},~\ref{fig:distinctiveness-detail}) to show how relationships between them reflect real-world behavior, this comparison becomes prohibitively cumbersome at larger timescales.  To visualize the time-varying high-dimensional content in a more intuitive way~\citep{HeusEtal18a} we projected the topic proportions matrices onto a two-dimensional space using Uniform Manifold Approximation and Projection~\citep[UMAP; ][]{McInEtal18}.  In this lower-dimensional space, each point represents a single video or recall event, and the distances between the points reflect the distances between the events' associated topic vectors (Fig.~\ref{fig:trajectory}). In other words, events that are nearer to each other in this space are more semantically similar, and those that are farther apart are less so.

@@

%\subsection*{Neurosynth decoding analyses}
%To gain further insights into the results from our video searchlight analysis, we constructed a voxelwise map of $p$-values by assigning each voxel the proportion of phase-shifted video trajectories to which its corresponding cube's timeseries was more highly correlated than the timeseries of the ``real" video.  This $p$-maps described the sensitivity of each voxel (cube) to the specific temporal structure of the video.  We then passed this $p$-map to \texttt{Neurosynth}.  Neurosynth does cool stuff.

\bibliography{../../CDL-bibliography/memlab}
% \bibliography{CDL-bibliography/memlab}

>>>>

\title{Memory for television episodes preserves event content while introducing new across-event similarities}
% I think this title still needs work. In this version of the paper, we don’t really focus on the across-event similarities. Let's revisit this once the edits are complete - AH

% Constructing mnemonic events from meaningfully structured experiences
% Transforming meaningfully structured experiences into mnemonic events

@@

\begin{abstract}
Our experiences unfold over time defining unique \textit{trajectories} through the relevant representational spaces.  By casting our life events as temporally evolving trajectories, one can compare the shape of the trajectory formed by an experience to that defined by our later remembering of that experience.  We propose a framework for mapping naturalistic experiences onto geometric spaces that characterize how experiences are segmented into discrete events, and how the contents of event sequences evolve over time.  We apply this approach to a naturalistic memory experiment which had participants view and recount a television episode.  The content of participants' recountings of events from the original episode closely matched the original episode's content. Further, we introduce two novel metrics for assessing memory quality (precision and distinctiveness), both of which relate to participants' ability to recapitulate the experience.  Lastly, we identified a network of brain structures that are sensitive to the ``shapes'' of ongoing experiences, and an overlapping network that is sensitive (at the time of encoding) to how people later remembered those experiences in relation to other experiences.  In this way, modeling the content of richly structured experiences can reveal how (geometrically and conceptually) those experiences are segmented into events and integrated into our memories of other experiences.
\end{abstract}

@@

Occasionally, this gradual ``drift" of our ongoing experience is punctuated by sudden changes, or ``shifts"~\citep[e.g., when we walk through a doorway; ][]{RadvZack17}.  Prior research suggests that these sharp transitions (termed \textit{event boundaries}) help to discretize our experiences (and their mental representations) into \textit{events}~\citep{RadvZack17, BrunEtal18, HeusEtal18b, ClewDava17, EzzyDava11, DuBrDava13}.  The interplay between the stable (within-event) and transient (across-event) temporal dynamics of an experience also provides a potential framework for transforming experiences into memories that distill those experiences down to their essence.  For example, prior work has shown that event boundaries can influence how we learn sequences of items~\citep{HeusEtal18b, DuBrDava13}, navigate~\citep{BrunEtal18}, and remember and understand narratives~\citep{ZwaaRadv98, EzzyDava11}. Prior research has implicated a network of brain regions (including the hippocampus and the medial prefrontal cortex) as playing a critical role in transforming experiences into structured and consolidated memories ~\citep{TompDava17}.

@@

\section*{Results}
To characterize the temporally dynamic contents of the \textit{Sherlock} episode and participants' subsequent recountings, we used a topic model~\citep{BleiEtal03} to discover the latent themes.  Topic models take as inputs a vocabulary of words to consider and a collection of text documents, and return two output matrices.  The first of these is a \textit{topics matrix} whose rows are topics (latent themes) and whose columns correspond to words in the vocabulary. The entries of the topics matrix define how each word in the vocabulary is weighted by each discovered topic.  For example, a detective-themed topic might weight heavily on words like ``crime,'' and ``search.''  The second output is a \textit{topic proportions matrix}, with one row per document and one column per topic.  The topic proportions matrix describes what mixture of discovered topics is reflected in each document.

\cite{ChenEtal17} collected hand-annotated information about each of 1000 (manually identified) time segments spanning the roughly 50 minute video used in their experiment.  This information included: a brief narrative description of what was happening, the location where the scene took place, the names of any characters on the screen, and other similar details (for a full list of annotated features, see \textit{Methods}).  We took from these annotations the union of all unique words (excluding stop words, such as ``and," ``or," ``but," etc.) across all features and scenes as the ``vocabulary" for the topic model.  We then concatenated the sets of words across all features contained in overlapping, sliding windows of (up to) 50 scenes, and treated each window as a single ``document" for the purpose of fitting the topic model.  Next, we fit a topic model with (up to) $K = 100$ topics to this collection of documents.  We found that 32 unique topics (with non-zero weights) were sufficient to describe the time-varying content of the video (see \textit{Methods}; Figs.~\ref{fig:schematic}, \topics).  Note that our approach is similar in some respects to Dynamic Topic Models~\citep{BleiLaff06} in that we sought to characterize how the thematic content of the episode evolved over time.  However, whereas Dynamic Topic Models are designed to characterize how the properties of \textit{collections} of documents change over time, our sliding window approach allows us to examine the topic dynamics within a single document (or video).  Specifically, our approach yielded (via the topic proportions matrix) a single \textit{topic vector} for each sliding window of annotations transformed by the topic model.  We then stretched (interpolated) the resulting windows-by-topics matrix to match the time series of the 1976 fMRI volumes collected as participants viewed the episode.

\begin{figure}[tp]
\centering
\includegraphics[width=1\textwidth]{figs/schematic}
\caption{\small \textbf{Methods overview.} We used hand-annotated descriptions of each moment of video to fit a topic model.  Three example video frames and their associated descriptions are displayed (top two rows).  Participants later recalled the video (in the third row, we show example recalls of the same three scenes from participant 17).  We used the topic model (fit to the annotations) to estimate topic vectors for each moment of video and each sentence the participants recalled.  Example topic vectors are displayed in the bottom row (blue: video annotations; green: example participant's recalls).  Three topic dimensions are shown (the highest-weighted topics for each of the three example scenes, respectively).  We also show the 10 highest-weighted words for each topic.  Figure~\topics~provides a full list of the top 10 words from each of the discovered topics.}
\label{fig:schematic}
\end{figure}
% Odd name for figure in "Results" section rather than "methods" section
% - Paxton
% The 10 top words in this figure seem to be out of order or incorrect. Paxton, can you look into this? - Andy

@@

\begin{figure}[tp]
\centering
\includegraphics[width=\textwidth]{figs/eventseg}
\caption{\small \textbf{Modelling naturalistic stimuli and recalls.} All panels: darker colors indicate greater values; range: [0, 1].  \textbf{A.} Topic vectors ($K = 100$) for each of the 1976 video timepoints.  \textbf{B.} Timepoint-by-timepoint correlation matrix of the topic vectors displayed in Panel A.  Event boundaries discovered by the HMM are denoted in yellow (30 events detected).  \textbf{C.} Average topic vectors for each of the 30 video events. \textbf{D.} Topic vectors for each of 265 sliding windows of sentences spoken by an example participant while recalling the video.  \textbf{E.} Timepoint-by-timepoint correlation matrix of the topic vectors displayed in Panel D. Event boundaries detected by the HMM are denoted in yellow (22 events detected).  For similar plots for all participants see Figure~\corrmats.  \textbf{F.} Average topic vectors for each of the 22 recalled events from the example participant.  \textbf{G.} Correlations between the topic vectors for every pair of video events (Panel C) and recalled events (from the example participant; Panel F).  For similar plots for all participants, see Figure~\matchmats.  \textbf{H.} Average correlations between each pair of video events and recalled events (across all 17 participants).  To create the figure, each recalled event was assigned to the video event with the most correlated topic vector (yellow boxes in panels G and H).}
\label{fig:model}
\end{figure}
% I'm all for citing software packages, but where what's the distinction between packages we do cite (seaborn, brainiak, sklearn, hypertools, quail, umap, etc.) and don't cite (wordcloud, nilearn, fastdtw, etc.)?
% - Paxton
% I also flagged the seaborn reference for removal here. Def not critical to understanding the figure - Andy

@@

Although the example participant's recall topic proportions matrix has some visual similarity to the video topic proportions matrix, the time-varying topic proportions for the example participant's recalls are not as sparse as those for the video (compare Figs.~\ref{fig:model}A and D).  Similarly, although there do appear to be periods of stability in the recall topic dynamics (i.e., most topics are active or inactive over contiguous blocks of time), the individual topics' overall timecourses are not as cleanly delineated as the video topics'.  To examine these patterns in detail, we computed the timepoint-by-timepoint correlation matrix for the example participant's recall topic trajectory (Fig.~\ref{fig:model}E).  As in the video correlation matrix (Fig.~\ref{fig:model}B), the example participant's recall correlation matrix has a strong block diagonal structure, indicating that their recalls are discretized into separated events.  As for the video correlation matrix, we leveraged an HMM-based optimization procedure (see \textit{Methods}) to determine how many events are reflected in the participant's recalls and where specifically the event boundaries fall (outlined in yellow).  We carried out a similar analysis on all 17 participants' recall topic proportions matrices (Fig.~\corrmats).

Two clear patterns emerged from this set of analyses.  First, although every individual participant's recalls could be segmented into discrete events (i.e., every individual participant's recall correlation matrix exhibited clear block diagonal structure; Fig.~\corrmats), each participant appeared to have a unique \textit{recall resolution}, reflected in the sizes of those blocks.  While, some participants' recall topic proportions segmented into just a few events (e.g., Participants P4, P5, and P7), others' segmented into many shorter duration events (e.g., Participants P12, P13, and P17).  This suggests that different participants may be recalling the video with different levels of detail-- e.g., some might touch on just the major plot points, whereas others might attempt to recall every minor scene or action.  The second clear pattern present in every individual participant's recall correlation matrix is that, unlike in the video correlation matrix, there are substantial off-diagonal correlations.  Whereas each event in the original video was (largely) separable from the others (Fig.~\ref{fig:model}B), in transforming those separable events into memory, participants appear to be integrating across multiple events, blending elements of previously recalled and not-yet-recalled content into each newly recalled event~\citep[Figs.~\ref{fig:model}E, \corrmats; also see][]{MannEtal11, HowaEtal12}.

The above results indicate that both the structure of the original video and participants' recalls of the video exhibit event boundaries that can be identified automatically by characterizing the dynamic content using a shared topic model and segmenting the content into events via HMMs.  Next, we asked whether some correspondence might be made between the specific content of the events the participants experienced in the video, and the events they later recalled.  One approach to linking the experienced (video) and recalled events is to label each recalled event as matching the video event with the most similar (i.e., most highly correlated) topic vector (Figs.~\ref{fig:model}G, \matchmats).  This yields a sequence of ``presented" events from the original video, and a (potentially differently ordered) sequence of ``recalled" events for each participant.  Analogous to classic list-learning studies, we can then examine participants' recall sequences by asking which events they tended to recall first~\citep[probability of first recall; Fig.~\ref{fig:list-learning}A;][]{AtkiShif68, PostPhil65, WelcBurn24}; how participants most often transition between recalls of the events as a function of the temporal distance between them~\citep[lag-conditional response probability; Fig.~\ref{fig:list-learning}B;][]{Kaha96}; and which events they were likely to remember overall~\citep[serial position recall analyses; Fig.~\ref{fig:list-learning}C;][]{Murd62a}. Interestingly, for two of these analyses (probability of first recall and lag-conditional response probability curves) we observe patterns comparable to classic effects from the list-learning literature: namely, a higher probability of initiating recall with the first event in the sequence (Fig.~\ref{fig:list-learning}A) and a higher probability of transitioning to neighboring events with an asymmetric forward bias (Fig.~\ref{fig:list-learning}B). In contrast, we do not observe a pattern comparable to the serial position effect (Fig.~\ref{fig:list-learning}C), but rather we see higher memory for specific events distributed somewhat evenly throughout the video.

We can also apply two list-learning-native analyses that describe how participants group items in their recall sequences: temporal clustering and semantic clustering \citep[][see \textit{Methods} for details]{PolyEtal09}.  Temporal clustering refers to the extent to which participants group their recall responses according to encoding position.  Overall, we found that sequentially viewed video events were clustered heavily in participants' recall event sequences (mean clustering score: 0.767, SEM: 0.029), and that participants with higher temporal clustering scores tended to perform better according to both \cite{ChenEtal17}'s hand-annotated memory scores (Pearson's $r(15) = 0.62,~p = 0.008$) and our model's estimate (Pearson's $r(15) = 0.54,~p = 0.024$).  Semantic clustering measures the extent to which participants cluster their recall responses according to semantic similarity.  We found that participants tended to recall semantically similar video events together (mean clustering score: 0.787, SEM: 0.018), and that semantic clustering score was also related to both hand-annotated  (Pearson's $r(15) = 0.65,~p = 0.004$) and model-derived (Pearson's $r(15) = 0.63,~p = 0.007$) memory performance.
% NOTE:
% Semantic and temporal clustering are both positively correlated with n_events -- r=0.49; p=0.046 and r=0.61; p=0.009, respectively.
% - Paxton

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=1\textwidth]{figs/precision_distinctiveness}
  \caption{\small \textbf{Novel content-based metrics of naturalistic memory: precision and distinctiveness.} \textbf{A.} The video-recall correlation matrix for a representative participant (17).  The yellow boxes highlight the maximum correlation in each column.  The example participant's overall precision score was computed as the average across correlation values in the yellow boxes.  Their distinctiveness score was computed as the average (over recall events) of 1 minus the average correlation between each recall event and all other recall events that do not display a box in the same row.  \textbf{B.} The (Pearson's) correlation between precision and hand-annotated memory performance. \textbf{C.} The correlation between distinctiveness and hand-annotated memory performance. \textbf{D.} The correlation between precision and the number of video events successfully recalled, as determined by our model. \textbf{E.} The correlation between distinctiveness and the number of video events successfully recalled, as determined by our model.}
  \label{fig:precision-distinctiveness}
\end{figure}
% Need to reconcile the description of distinctiveness with how we computed it... - Andy

Statistical models of memory studies often treat recall success as binary (i.e., an item either was or was not recalled), or occasionally categorical \citep[e.g., to distinguish familiarity from recollection;][]{YoneEtal02}.  Such approaches are tenable in classical list-learning or recognition memory paradigms, as the presented stimuli tend to be very simple (e.g., a sequence of individual words or items).  However, the feature-rich content of a naturalistic experiences may later be described with many, highly variable levels of success.  Our framework produces a content-based model of individual stimulus and recall events by projecting the dynamic content of the video and participants' recalls into a shared topic space.  This allows for direct, quantitative comparison between all stimulus and recall events, as well as between the recall events themselves.  Leveraging these content-based models of the stimulus/recall events, we developed two novel, \textit{continuous} metrics for analyzing naturalistic memory:~\textit{precision} and \textit{distinctiveness}.  We define precision as the ``completeness" of recall, or how fully the presented content was recapitulated in memory.  Under our framework, we quantify this for a given recall event as the correlation between the topic proportions of the recall event and the maximally correlated video event (Fig.~\ref{fig:precision-distinctiveness}).  A second novel metric we introduce here is \textit{distinctiveness}, which we define as the ``specificity" of recall, or how unique the description of a given section of content was, compared to descriptions for other sections of content.  We quantify this for each recall event as 1 minus the average correlation between the given recall event and all other recall events not matched to the same video event.  In addition to individual events, one may also use these metrics to describe each participant's overall performance (i.e., by averaging across a participant's event-wise precision or distinctiveness scores).  Participants whose recall events are more veridical descriptions of what happened in the video event will presumably have higher precision scores. We find that, across participants, a higher precision score is correlated to both hand-annotated memory performance (Pearson's $r(15) = 0.60, p = 0.010$) and the number of video events successfully remembered, as determined by our model (Pearson's $r(15) = 0.90, p < 0.001$).  We also hypothesized that participants who recounted events in a more distinctive way would display better overall memory.  We find that this distinctiveness score is related to our model's estimated number of recalled events (Pearson's $r(15) = 0.55, p = 0.021$), and while we do not find distinctiveness to be related to hand-annotated memory performance (Pearson's $r(15) = 0.28, p = 0.275$), this is not entirely surprising given how the hand-annotated memory scores were computed (see \textit{Methods} and \textit{Discussion} for details).

@@

Figure \ref{fig:distinctiveness-detail} similarly contrasts two example participants' recalls for a common video event (event 19) to illustrate the tangible differences between high and low distinctiveness scores.  Here, we have extracted the full set of sentences comprising the most distinctive recall event (P13) and least distinctive recall event (P11) matched to the example video event (Fig.~\ref{fig:distinctiveness-detail}C).  We also extracted the annotations for the example video event, as well as those from each other video event whose content the example participants' single recall events described (Fig.~\ref{fig:distinctiveness-detail}B).  We then shaded the annotation text for each video event with a different color, and shaded each word of the example participants' recall text by the color of the video event it describes.  The majority of the most distinctive recall event text describes video event 19's content, with the first five and last one sentence describing the video events immediately preceding and succeeding the current one, respectively.  In contrast, the least precise participant's recall for video event 19 blends the content from five separate video events, does not transition between them in order, and often combines descriptions of two video events' content in the same sentence.

The prior analyses leverage the correspondence between the 100-dimensional topic proportion matrices for the video and participants' recalls to characterize recall.  However, it is difficult to gain deep insights into the content of (or relationships between) experiences and memories solely by examining these topic proportions (e.g., Figs.~\ref{fig:model}A, D) or the corresponding correlation matrices (Figs.~\ref{fig:model}B, E, \corrmats).  And while we can directly examine the original text underlying these topic vectors (e.g., Figs.~\ref{fig:precision-detail},~\ref{fig:distinctiveness-detail}) to show how relationships between them reflect real-world behavior, this comparison becomes prohibitively cumbersome at larger timescales.  To visualize the time-varying high-dimensional content in a more intuitive way~\citep{HeusEtal18a}, we projected the topic proportions matrices onto a two-dimensional space using Uniform Manifold Approximation and Projection~\citep[UMAP; ][]{McInEtal18}.  In this lower-dimensional space, each point represents a single video or recall event, and the distances between the points reflect the distances between the events' associated topic vectors (Fig.~\ref{fig:trajectory}). In other words, events that are nearer to each other in this space are more semantically similar, and those that are farther apart are less so.

@@

\subsection*{Neurosynth decoding analyses}
\texttt{Neurosynth} parses a massive online database of over 14,000 neuroimaging studies and constructs meta-analysis images for over 13,000 psychology- and neuroscience-related terms, based on NIfTI images accompanying studies where those terms appear at a high frequency.  Then, given a novel image (tagged with its value type; e.g., $t$-, $F$- or $p$-statistics), \texttt{Neurosynth} returns a list of terms whose meta-analysis images are most similar to this new data.  Our permutation procedure yielded, for each of the two searchlight analyses, a voxelwise map of significance ($p$-statistic) values.  These maps describe the extent to which each voxel \textit{specifically} reflected the temporal structure of the video or individuals' recalls (i.e., for each voxel, the proportion of phase-shifted topic vector correlation matrices less similar to the voxel activity correlation matrix than the unshifted video's correlation matrix). We input the two statistical maps described above to \texttt{Neurosynth} to create a list of the 10 most representative terms for each map.
